Discrete Quantum Theories and Computing

### 2019-01-19 dissertation.tex
dissertation.lyx Update according to the suggestion from Spellchecker and Check TeX...

%% LyX 2.3.2-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{rotfloat}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged... }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:QIVPM} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM) \cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.
Dylan Thurston for serving in my advisory committee and my Tier 3
committee, with other advisory committee members passing my Computer
Science qualifying exam Spring 2016. Especially, my survey of real
computation in the qualifying exam inspired me the comparison of quantum
computer and analog computer in Chapter~\ref{chap:Introduction}.
Chapter~\ref{chap:Introduction} also contains the motivation in
my application for Rethinking Foundations of Physics 2017 workshop
reviewed by Prof. Amr Sabry(\textarabic{عمرو صبري}). I want
to thank Prof. Lawrence Moss for inviting me to present our results
in the interdisciplinary logic seminar and theory seminar in Computer
Science, co-chairing my Tier 3 committee, writing an assessment letter
for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth ``Betsy'' Merceron and Kexin ``Casey''
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

\todo{Type a table for acknowledgment so that I would have to write
``I thank'' so many times here? }Many people in both department
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof. Christopher
Judge and Prof. Matthias Weber; directors of PhD. study in Computer
Science, especially Prof. Yuqing Melanie Wu (吴愈青) and Prof. \textturkish{Funda
Ergun}.

Thank Kin Wai Chan for suggesting me the correct direction for Proposition~\ref{appH.sec},
Prof. 朱樺 for drawing my attention on Prof. Wan's book~\cite{Wan2006},
and Weihua Liu (刘伟华) for remaining me analysts call the right side
of the IVPM outer measure.

@@

\item My other AT\&T family plan members, especially Kexin ``Casey''
Chen (谌可心), Yao-yu Chih (池耀宇), Xinyi ``Diana'' Gong (龚欣怡), and Nanxi
``Nancy'' Zhang (张楠茜), for you always pay our telephone bill on
time so that......................\todo{Maybe I don't have to thank
them as AT\&T family plan members, but I should thank some of them
for support and sharing information...}

@@

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX ~\cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows with MiK\TeX ~2.9.
Please report me if there is any difficulty to compile it. Besides,
to compile a PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git},
open \code{dissertation.lyx} in \LyX , and click the toolbar button
\includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Similar to its classical counterparts, quantum circuit model,
the most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, taking into account the
precision of quantities could still result in a better model. However,
comparing to the precision of probability amplitudes, it is easier
to consider the precision of their inducing probabilities because
the idea of ``imprecise probabilities'' have well-studied classically
\cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities... }This idea will be extended to quantum
domain as quantum interval-valued probability measures (QIVPM).

@@

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably unphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers by a unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
It is called the modal quantum theory because it can only predict
whether a measure result is possible or impossible, but not its probability.
Although a version of quantum theory defined over the two-valued field
can express simple algorithms such as quantum teleportation \cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to efficiently solve the $\usat$ problem~\cite{Papadimitriou1993},
which is as hard as a general $\NP$-complete problem~\cite{Valiant198685}.

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to \ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that preserve a great many of the standard requirements
of quantum computing. In this model, the state of a discrete $d$-dimensional
system is a vector in the $d$-dimensional vector space over $\Fpp$,
and can be reduced to a point in the discrete complex projective space.
Among these irreducible states, the product states and entangled states
could not only be identified as in CQT but also be counted for different
$p$. Despite this model's similarity to CQT, it still only predicts
whether a measurement result is possible or impossible, because it
is not possible to define an inner product in the usual sense due
to the modular arithmetic~\cite{grove2002classical}. Under suitable
conditions, we can have deterministic quantum algorithms such as the
algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem efficiently.

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally Chapter~\ref{chap:Further-Discussion} contains the further
extended questions.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to a quantum probability measure in CQT. However, since the
variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, searching for a Born rule step-by-step here might provide
a better idea what we should do in Sec.~\ref{sec:Toward-IVPM} and
Chapter~\ref{chap:QIVPM}. While the situations will become more
delicate in the later sections, we will fortunately find the unique
Born rule for CQT here.

@@

\section{Discrete Quantum Computing (II)}

\label{discretequantumcomputingII}

@@

\todo{Add some introductory text...}

@@

\section{The Kochen-Specker Theorem and Contextuality}

\label{sec:Kochen-Specker}

>>>>

%% LyX 2.3.2-2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{rotfloat}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged\ldots{} }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:QIVPM} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM)~\cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.\ Dylan
Thurston for serving in my advisory committee and my Tier 3 committee,
with other advisory committee members passing my Computer Science
qualifying exam Spring 2016. Especially, my survey of real computation
in the qualifying exam inspired me the comparison of quantum computer
and analog computer in Chapter~\ref{chap:Introduction}. Chapter~\ref{chap:Introduction}
also contains the motivation in my application for Rethinking Foundations
of Physics 2017 workshop reviewed by Prof.\ Amr Sabry~(\textarabic{عمرو صبري}).
I want to thank Prof.\ Lawrence Moss for inviting me to present our
results in the interdisciplinary logic seminar and theory seminar
in Computer Science, co-chairing my Tier 3 committee, writing an assessment
letter for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth ``Betsy'' Merceron and Kexin ``Casey''
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

\todo{Type a table for acknowledgment so that I would have to write
``I thank'' so many times here? }Many people in both department
of Mathematics and Computer Science need to do something more to make
my double-major works smoothly. I need to thank all of them, especially
directors of graduate study in Mathematics, especially Prof.\ Christopher
Judge and Prof.\ Matthias Weber; directors of PhD. study in Computer
Science, especially Prof.\ Yuqing Melanie Wu (吴愈青) and Prof.\ \textturkish{Funda
Ergun}.

Thank Kin Wai Chan for suggesting me the correct direction for Proposition~\ref{appH.sec},
Prof.\ 朱樺 for drawing my attention on Prof.\ Wan's book~\cite{Wan2006},
and Weihua Liu (刘伟华) for remaining me analysts call the right side
of the IVPM outer measure.

@@

\item My other AT\&T family plan members, especially Kexin ``Casey''
Chen (谌可心), Yao-yu Chih (池耀宇), Xinyi ``Diana'' Gong (龚欣怡), and Nanxi
``Nancy'' Zhang (张楠茜), for you always pay our telephone bill on
time so that\ldots\ldots\todo{Maybe I don't have to thank them
as AT\&T family plan members, but I should thank some of them for
support and sharing information\ldots}

@@

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX{} \cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows with MiK\TeX{}
2.9. Please report me if there is any difficulty to compile it. Besides,
to compile a PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git},
open \code{dissertation.lyx} in \LyX , and click the toolbar button
\includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Similar to its classical counterparts, quantum circuit model,
the most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, taking into account the
precision of quantities could still result in a better model. However,
comparing to the precision of probability amplitudes, it is easier
to consider the precision of their inducing probabilities because
the idea of ``imprecise probabilities'' have well-studied classically
\cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }This idea will be extended
to quantum domain as quantum interval-valued probability measures
(QIVPM).

@@

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably nonphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers by a unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
It is called the modal quantum theory because it can only predict
whether a measure result is possible or impossible, but not its probability.
Although a version of quantum theory defined over the two-valued field
can express simple algorithms such as quantum teleportation \cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to efficiently solve the $\usat$ problem~\cite{Papadimitriou1993},
which is as hard as a general $\NP$-complete problem~\cite{Valiant198685}.

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to~\ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that preserve a great many of the standard requirements
of quantum computing. In this model, the state of a discrete $d$-dimensional
system is a vector in the $d$-dimensional vector space over $\Fpp$,
and can be reduced to a point in the discrete complex projective space.
Among these irreducible states, the product states and entangled states
could not only be identified as in CQT but also be counted for different
$p$. Despite this model's similarity to CQT, it still only predicts
whether a measurement result is possible or impossible, because it
is not possible to define an inner product in the usual sense due
to the modular arithmetic~\cite{grove2002classical}. Under suitable
conditions, we can have deterministic quantum algorithms such as the
algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem efficiently.

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to~\ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally Chapter~\ref{chap:Further-Discussion} contains the further
extended questions.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to a quantum probability measure in CQT\@. However, since
the variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, searching for a Born rule step-by-step here might provide
a better idea what we should do in Sec.~\ref{sec:Toward-IVPM} and
Chapter~\ref{chap:QIVPM}. While the situations will become more
delicate in the later sections, we will fortunately find the unique
Born rule for CQT here.

@@

\section{Discrete Quantum Computing (II)\label{discretequantumcomputingII}}

@@

\todo{Add some introductory text\ldots}

@@

\section{The Kochen-Specker Theorem and Contextuality\label{sec:Kochen-Specker}}
